---
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE, 
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```


# ggFishPlots
**Visualise and calculate life history parameters for fisheries science using ggplot2. R package version `r packageVersion("ggFishPlots")`**

<!-- badges: start -->
[![R-CMD-check](https://github.com/DeepWaterIMR/ggFishPlots/workflows/R-CMD-check/badge.svg)](https://github.com/DeepWaterIMR/ggFishPlots/actions)
<!-- badges: end -->

## Overview

The ggFishPlots package for [R](https://www.r-project.org/) allows quickly plotting and calculating life history parameters required by stock assessment models. The package uses [ggplot2](https://ggplot2.tidyverse.org/reference) for plotting and the [tidyverse](https://www.tidyverse.org/) packages for calculations.

The ggOceanMaps package has been developed by the [Institute of Marine Research](https://www.hi.no/en). Note that the package comes with absolutely no warranty. Any [bug reports and code fixes](https://github.com/DeepWaterIMR/ggFishPlots/issues) are warmly welcomed. See [*Contributions*](#contributions) for further details.

If you are looking for other similar packages, make sure to check out  [AquaticLifeHistory](https://jonathansmart.github.io/AquaticLifeHistory/articles/Maturity_analyses.html).

## Installation

The package is currently available only on [GitHub](https://github.com/DeepWaterIMR/ggFishPlots). It can be installed directly using the [**devtools**](https://cran.r-project.org/web/packages/devtools/index.html) or remotes packages. 

```{r eval = FALSE}
devtools::install_github("DeepWaterIMR/ggFishPlots")
```

## Usage

At the time of writing, the package produces three kinds of plots: growth curves, maturity plots, and length-weight relationships. Each function returns a ggplot2 plot and the estimated parameters as a text string that can be used in Rmarkdown and Shiny applications as well as a data frame for further use of the parameters. The elements are returned as a list. The package contains example data to illustrate the functionality.

### Growth curves

Simple plot. Note how the `text` and `params` are returned as a list together with `plot`.

```{r}
library(ggFishPlots)

data(survey_ghl) # example data

head(survey_ghl)

plot_growth(survey_ghl, length = "length", age = "age")
```

Split by sex. Specifying `length`, `age` and `sex` arguments have been omitted since they are the argument names by default and the same than in the example data. Only the `plot` element is returned this time.

```{r}
plot_growth(survey_ghl, split.by.sex = TRUE)$plot
```

The dashed lines are $S_{inf}$. Data behind the growth curves are shown as box plots by default. It is possible to plot the data as points by defining `boxplot = FALSE`. We can also force zero group into the curves if know the length of it. Assumed as 10 cm here: 

```{r}
plot_growth(survey_ghl, force.zero.group.length = 10, boxplot = FALSE)$plot
```

### Maturity plots

Simple L50 plot:

```{r}
plot_maturity(survey_ghl, length = "length", maturity = "maturity")
```

The error bars represent 95% confidence intervals calculated from the model object using the [`confint()`](https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/confint.html) function and back-transformed to the original scale. The grey stepped line is a binned average defined using the `length.bin.width` argument.

Split by sex:

```{r}
plot_maturity(survey_ghl, split.by.sex = TRUE)$plot
```

The same principle can be used to produce A50 (50% maturity at age) plots:

```{r}
plot_maturity(survey_ghl, length = "age", length.unit = "years",
              xlab = "Age", length.bin.width = 1, split.by.sex = TRUE)$plot
```

### Length-weight relationships

Simple plot using log-transformation and linear models by default:

```{r}
plot_lw(survey_ghl, length = "length", weight = "weight")
```

The dashed lines represent 95% confidence intervals. Use non-linear least squares instead:

```{r}
plot_lw(survey_ghl, use.nls = TRUE)
```

Split by sex:

```{r}
plot_lw(survey_ghl, split.by.sex = TRUE)$plot
```

Log-log axes to see differences:

```{r}
plot_lw(survey_ghl, split.by.sex = TRUE, log.axes = TRUE)$plot
```

Note that the floating point in the parameters depends on the [length and weight units](https://www.fishbase.de/manual/fishbasethe_length_weight_table.htm). According to the FishBase, one should use centimeters and grams but you can also transform the parameters according to the formulas given in the FishBase too.

```{r}
plot_lw(survey_ghl %>% dplyr::mutate(weight = weight*1000), weight.unit = "g")$params
```

## Citations and data sources

The data used in the package are a property of the Institute of Marine Research and the Norwegian Government. They are distributed under the Creative Commons ([CCBY](https://creativecommons.org/licenses/by/4.0/) or [NLOD](https://data.norge.no/nlod/no/1.0/)) licenses allowing free use as long as the source (IMR) is cited. We ask any user to refer to the package if plots or estimates are used in reports or scientific articles. For up-to-date citation information, please use:

```{r}
citation("ggFishPlots")
```

## Contributions

Any contributions to the package are more than welcome. Please contact the package maintainer Mikko Vihtakari (<mikko.vihtakari@hi.no>) to discuss your ideas on improving the package. Bug reports and corrections should be submitted directly to [the GitHub site](https://github.com/DeepWaterIMR/ggFishPlots/issues). Please include a [minimal reproducible example](https://en.wikipedia.org/wiki/Minimal_working_example). Considerable contributions to the package development will be credited with an authorship. 
