---
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE, 
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```


# ggFishPlots
**Visualise and calculate life history parameters for fisheries science using ggplot2. R package version `r packageVersion("ggFishPlots")`**

<!-- badges: start -->
[![R-CMD-check](https://github.com/DeepWaterIMR/ggFishPlots/workflows/R-CMD-check/badge.svg)](https://github.com/DeepWaterIMR/ggFishPlots/actions)
<!-- badges: end -->

## Overview

The ggFishPlots package for [R](https://www.r-project.org/) allows quickly plotting and calculating life history parameters required by stock assessment models from data. The package uses [ggplot2](https://ggplot2.tidyverse.org/reference) for plotting and the [tidyverse](https://www.tidyverse.org/) packages for calculations.

The ggOceanMaps package has been developed by the [Institute of Marine Research](https://www.hi.no/en). Note that the package comes with absolutely no warranty. Any [bug reports and code fixes](https://github.com/DeepWaterIMR/ggFishPlots/issues) are warmly welcomed. See [*Contributions*](#contributions) for further details.

If you are looking for other similar packages, make sure to check out the [AquaticLifeHistory](https://jonathansmart.github.io/AquaticLifeHistory/articles/Maturity_analyses.html).

## Installation

The package is currently available only on [GitHub](https://github.com/DeepWaterIMR/ggFishPlots). It can be installed directly using the [**devtools**](https://cran.r-project.org/web/packages/devtools/index.html) or remotes packages. 

```{r eval = FALSE}
devtools::install_github("DeepWaterIMR/ggFishPlots")
```

## Usage

At the time of writing, the package produces three kinds of plots: growth curves, maturity plots, and length-weight relationships. Each function returns a ggplot2 plot and the estimated parameters as a list. The package contains example data to illustrate the functionality.

### Growth curves

Simple plot. Note how the `text` and `params` are returned as a list together with `plot`.

```{r}
library(ggFishPlots)

data(survey_ghl) # example data

plot_growth(survey_ghl, length = "length", age = "age")
```

Split by sex. Specifying `length`, `age` and `sex` arguments omitted since they are the argument names by default and the same than in the example data. Only the `plot` element returned this time.

```{r}
plot_growth(survey_ghl, split.by.sex = TRUE)$plot
```

The dashed lines are $S_{inf}$. Data behind the growth curves are shown as boxplots by default. It is possible to plot the data as points by defining `boxplot = FALSE`. We can also force zero group into the curves if know the length of it. Assumed as 10 cm here: 

```{r}
plot_growth(survey_ghl, force.zero.group.length = 10, boxplot = FALSE)$plot
```

### Maturity plots

Simple L50 plot:

```{r}
plot_maturity(survey_ghl, length = "length", maturity = "maturity")
```

Split by sex:

```{r}
plot_maturity(survey_ghl, split.by.sex = TRUE)$plot
```

The same principle can be used to produce A50 (50% maturity at age) plots:

```{r}
plot_maturity(survey_ghl, length = "age", length.unit = "years",
              xlab = "Age", length.bin.width = 1, split.by.sex = TRUE)$plot
```

### Length-weight relationships

Simple plot using log-transformation and linear models by default:

```{r}
plot_lw(survey_ghl, length = "length", weight = "weight")
```

Use non-linear least squares instead:

```{r}
plot_lw(survey_ghl, use.nls = TRUE)
```

Split by sex:

```{r}
plot_lw(survey_ghl, split.by.sex = TRUE)$plot
```

Log-log axes to see differences:

```{r}
plot_lw(survey_ghl, split.by.sex = TRUE, log.axes = TRUE)$plot
```


Note that the floating point in the parameters depends on the [length and weight units](https://www.fishbase.de/manual/fishbasethe_length_weight_table.htm). According to FishBase one should use centimeters and grams but you can also transform the parameters according to the formulas given in FishBase too.

```{r}
plot_lw(survey_ghl %>% dplyr::mutate(weight = weight*1000), weight.unit = "g")$params
```

## Citations and data sources

The data used in the package is property of the Institute of Marine Research and are open domain as long as the source (IMR) is cited. We ask any user to refer to the package if plots or estimates are used in reports or scientific articles. For up-to-date citation information, please use:

```{r}
citation("ggFishPlots")
```

## Contributions

Any contributions to the package are more than welcome. Please contact the package maintainer Mikko Vihtakari (<mikko.vihtakari@hi.no>) to discuss your ideas on improving the package. Bug reports and corrections should be submitted directly to [the GitHub site](https://github.com/DeepWaterIMR/ggFishPlots/issues). Please include a [minimal reproducible example](https://en.wikipedia.org/wiki/Minimal_working_example). Considerable contributions to the package development will be credited with an authorship. 
