---
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE, 
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```


# ggFishPlots
**Visualise and calculate life history parameters for fisheries science using ggplot2. R package version `r packageVersion("ggFishPlots")`**

<!-- badges: start -->
[![DOI](https://zenodo.org/badge/DOI/10.32614/CRAN.package.ggFishPlots.svg)](https://doi.org/10.32614/CRAN.package.ggFishPlots)
[![R-CMD-check](https://github.com/DeepWaterIMR/ggFishPlots/workflows/R-CMD-check/badge.svg)](https://github.com/DeepWaterIMR/ggFishPlots/actions)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/ggFishPlots)](https://CRAN.R-project.org/package=ggFishPlots)
<!-- badges: end -->

## Overview

The [ggFishPlots](https://deepwaterimr.github.io/ggFishPlots/) package for [R](https://www.r-project.org/) allows quickly plotting and calculating life history parameters required by stock assessment models. The package uses [ggplot2](https://ggplot2.tidyverse.org/reference) for plotting and the [tidyverse](https://www.tidyverse.org/) packages for calculations.

The ggOceanMaps package has been developed by the [Institute of Marine Research](https://www.hi.no/en). Note that the package comes with absolutely no warranty. Any [bug reports and code fixes](https://github.com/DeepWaterIMR/ggFishPlots/issues) are warmly welcomed. See [*Contributions*](#contributions) for further details.

If you are looking for other similar packages, make sure to check out the [AquaticLifeHistory](https://jonathansmart.github.io/AquaticLifeHistory/articles/Maturity_analyses.html),  [FSA](https://CRAN.R-project.org/package=FSA), and [fishmethods](https://CRAN.R-project.org/package=fishmethods) packages, as well as the [FishR webpage](https://fishr-core-team.github.io/fishR/pages/packages.html).

## Installation

The most up to date version of the package can always be found from [GitHub](https://github.com/DeepWaterIMR/ggFishPlots). If the [CRAN version](https://CRAN.R-project.org/package=ggFishPlots) does not work as shown in the examples on this website, try installing the GitHub version. You can do this by using the [devtools](https://cran.r-project.org/web/packages/devtools/index.html) or [remotes](https://cran.r-project.org/web/packages/remotes/index.html) packages. 

```{r eval = FALSE}
remotes::install_github("DeepWaterIMR/ggFishPlots")
```

## Usage

The package has a [website](https://deepwaterimr.github.io/ggFishPlots/). At the time of writing, the package produces four kinds of plots: [growth curves](#growth-curves), [maturity plots](#maturity-plots), [length-weight relationships](#length-weight-relationships), and [catch curves](#catch-curves-to-estimate-instantaneous-total-mortality-z). Each function returns a ggplot2 plot and the estimated parameters as a text string that can be used in Rmarkdown and Shiny applications as well as a data frame for further use of the parameters. The elements are returned as a list. The package contains example data to illustrate the functionality.

### Growth curves

Note how the `text` and `params` are returned as a list together with `plot`.

```{r}
library(ggFishPlots)

data(survey_ghl) # example data

head(survey_ghl)

plot_growth(survey_ghl, length = "length", age = "age")
```

The `text` object can be rendered to R markdown documents using the `results = 'asis'` setting in the code chunk header (i.e. `{r, results = 'asis'}`) and the `cat()` function after replacing `"\n"` by `"\  n"`:

```{r, results='asis'}
htmlcat <- function(text){
  cat(gsub(pattern = "\n", replacement = "  \n", x = text))
}

htmlcat(plot_growth(survey_ghl)$text)
```

#### Split by sex

Specifying `length`, `age` and `sex` arguments have been omitted since they are the argument names by default and the same than in the example data. Only the `plot` element is returned this time.

```{r}
plot_growth(survey_ghl, split.by.sex = TRUE)$plot
```

The dashed lines are $S_{inf}$. Data behind the growth curves are shown as box plots by default. It is possible to plot the data as points by defining `boxplot = FALSE`. We can also force zero group into the curves if know the length of it. Assumed as 14 cm here. Zero group forcing strength is 10 % of number of observations by default and can be adjusted using the `force.zero.group.strength` argument.

```{r}
plot_growth(survey_ghl, force.zero.group.length = 14, boxplot = FALSE)$plot
```

### Maturity plots

#### L50 plots

Maturity ogives are estimated using a logistic (`family = binomial(link = "logit")`) [`glm()`](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/glm)

```{r}
plot_maturity(survey_ghl, length = "length", maturity = "maturity")
```

The error bars represent 95% confidence intervals calculated from the model object using the [`confint()`](https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/confint.html) function and back-transformed to the original scale. The grey stepped line is a binned average defined using the `length.bin.width` argument. 

The function also contains an option to bootstrap the confidence intervals (CIs) which will produce narrower CIs. Bootstrapping is probably a more correct way of estimating CIs in this application. Using only 10 replicates here to save processing time. In real application use at least 1000.

```{r}
plot_maturity(survey_ghl, bootstrap.n = 10)
```

#### Split by sex

```{r}
plot_maturity(survey_ghl, split.by.sex = TRUE)$plot
```

#### A50 plots

The same principle can be used to produce A50 (50% maturity at age) plots:

```{r}
plot_maturity(survey_ghl, length = "age", length.unit = "years",
              xlab = "Age", length.bin.width = 1, split.by.sex = TRUE)$plot
```

#### Addition of juveniles

Also the `plot_maturity()` function has the option to add juveniles (zero group fish). The addition of juveniles can be necessary for to make the `glm()` to converge if small immature fish are scarce in the dataset. 

```{r}
plot_maturity(survey_ghl, length = "age", length.unit = "years",
              xlab = "Age", length.bin.width = 1, 
              force.zero.group.length = 0,
              force.zero.group.strength = 100,
              split.by.sex = TRUE)$plot
```

Note how the addition increases the estimate for males because the dataset did not contain a sufficient number of young males which the number of females seems sufficient because females mature older in this species.

One option can be to estimate the number of added juveniles using the [`plot_catchcurve()`](#catchcurvesep) function:

```{r}
plot_maturity(survey_ghl, length = "age", length.unit = "years",
              xlab = "Age", length.bin.width = 1, 
              force.zero.group.length = 0,
              force.zero.group.n = c("F" = exp(11.363), "M" = exp(11.885)),
              split.by.sex = TRUE)$plot
```

### Length-weight relationships

#### Log-linearized models

Simple plot using log-transformation and linear models by default.

```{r}
plot_lw(survey_ghl, length = "length", weight = "weight")
```

The dashed lines represent 95% confidence intervals. 

#### Non-linear least square models

Use non-linear least squares instead:

```{r}
plot_lw(survey_ghl, use.nls = TRUE)
```

#### Split by sex

The decimal point of a and b estimates depends on the length and weight units. [FishBase](https://www.fishbase.se/manual/english/FishBaseThe_LENGTH_WEIGHT_Table.htm) uses centimeters and grams. The function can correct for the units when asked (but `length.unit` and `weight.unit` parameters have to be defined correctly).

```{r}
plot_lw(survey_ghl, split.by.sex = TRUE, correct.units = TRUE)
```

You can also transform the parameters according to the formulas given in the FishBase.

```{r}
plot_lw(survey_ghl %>% dplyr::mutate(weight = weight*1000), weight.unit = "g")$params
```

#### Log-log axes to see differences

```{r}
plot_lw(survey_ghl, split.by.sex = TRUE, log.axes = TRUE)$plot
```

#### Outlier removal 

```{r}
plot_lw(survey_ghl, outlier.percentile = 99.5, annotate.coefficients = TRUE)$plot
```

### Catch curves to estimate instantaneous total mortality (Z)

Catch curves were added to the version 0.2.3 and may not be available in the CRAN version. [A FishR tutorial written by Ogle (2013)](https://www.fishbase.se/manual/english/FishBaseThe_LENGTH_WEIGHT_Table.htm) contains a nice explanation of catch curves. At the time of writing ggFishPlots calculates only the simple log-linearised regression. 

```{r}
plot_catchcurve(survey_ghl)
```

The ages to be included to the Z estimation can be adjusted using the `age.range` argument.

```{r}
plot_catchcurve(survey_ghl, age.range = c(10,26))$plot
```

In the plot above, -b is Z (i.e. Z = 0.356) and exp(a) (i.e. 68391) is the number of 0 age fish assuming constant mortality.

#### Split by sex {#catchcurvesep}

```{r}
plot_catchcurve(survey_ghl, age.range = c(10,26), split.by.sex = TRUE)$plot
```

Use a named list to use separate age ranges for females and males.

```{r}
tmp <- plot_catchcurve(survey_ghl,
age.range = list("F" = c(13,26), "M" = c(10,26)),
split.by.sex = TRUE)

tmp$plot
```

Parameter estimates for the catch curve above:

```{r, results='asis'}
htmlcat(tmp$text)
```

## Citations and data sources

The data used in the package are a property of the Institute of Marine Research and the Norwegian Government. They are distributed under the Creative Commons ([CCBY](https://creativecommons.org/licenses/by/4.0/) or [NLOD](https://data.norge.no/nlod/no/1.0/)) licenses allowing free use as long as the source (IMR) is cited. We ask any user to refer to the package if plots or estimates are used in reports or scientific articles. For up-to-date citation information, please use:

```{r}
citation("ggFishPlots")
```

## Contributions

Any contributions to the package are more than welcome. Please contact the package maintainer Mikko Vihtakari (<mikko.vihtakari@hi.no>) to discuss your ideas on improving the package. Bug reports and corrections should be submitted directly to [the GitHub site](https://github.com/DeepWaterIMR/ggFishPlots/issues). Please include a [minimal reproducible example](https://en.wikipedia.org/wiki/Minimal_working_example). Considerable contributions to the package development will be credited with an authorship. 
